A Fast Unified Model for Parsing and Sentence Understanding

### 2016-05-31 writing/hard_stack_paper/paper.tex
Rewrite thin-stack intro

A na\"ive implementation of SPINN would require representing a stack of size $N$ for each timestep of each input sentence at training time to support backpropagation. This implies a per-example space requirement of $N \times T \times D$, which is prohibitively large for significant batch sizes or sentence lengths $N$. Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the stack.

We propose an alternative space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S[t]$ (for $0<t\le T$) represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. In order to perform the \reduce\ operation, we need to retrieve the top two elements of the actual stack. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. The top two elements of the stack can be found by using the final two pointers in the queue $Q$. These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.

>>>>

A na\"ive implementation of SPINN needs to handle a size $O(N)$ stack at each timestep in an example, any element of which may be involved in later computations. A na\"ive backpropagation implementation would then require storing each of the $O(N)$ stacks for a backward pass, leading to a per-example space requirement of $O(NTD)$ floats. This requirement is prohibitively large for significant batch sizes or sentence lengths $N$. Such a na\"ive implementation would also require copying a largely unchanged stack at each timestep, since each \shift\ or \reduce\ operation writes only one new representation to the top of the stack.

We propose a space-efficient stack representation inspired by the zipper technique \citep{huet1997zipper}, that we call \textit{thin stack}. For each input sentence, we represent the stack with a single $T \times D$ matrix $S$. Each row $S[t]$ (for $0<t\le T$) represents the top of the actual stack at timestep $t$. At each timestep we can \shift\ a new element onto the stack, or \reduce\ the top two elements of the stack into a single element. To shift an element from the buffer to the top of the stack at timestep $t$, we simply write it into the location $S[t]$. In order to perform the \reduce\ operation, we need to retrieve the top two elements of the actual stack. We maintain a queue $Q$ of pointers into $S$ which contains the row indices of $S$ which are still present in the actual stack. The top two elements of the stack can be found by using the final two pointers in the queue $Q$. These retrieved rows are used to perform the \reduce\ operation, which modifies $Q$ to mark that some rows of $S$ have now been replaced in the actual stack. Algorithm~\ref{alg:thin-stack} describes the full mechanics of a stack feedforward in this compressed representation. It operates on the single $T \times D$ matrix $S$ and a backpointer queue $Q$. Table~\ref{tbl:thin-stack} shows an example run.
